---
title: "Loan Prediction"
author: "Pravesh Agrawal"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the required package
```{r}
pacman::p_load(caret)
```

## Reading the Train Data Set
```{r}
train <- read.csv("train_ctrUa4K.csv")
```

## Finding the unique values
```{r}
sapply(train,function(x)length(unique(x)))
sapply(train[,c(2,3,4,5,6,10,11,12,13)], function(x)unique(x))
```

## Replacing the blank fields with NA
```{r}
sapply(train,function(x)sum(is.na(x)))


train$Gender <- factor(train$Gender,levels = c("Male","Female"), labels = c("Male","Female"))

train$Married <- factor(train$Married,levels = c("No","Yes"), labels = c("No","Yes"))

train$Dependents <- factor(train$Dependents,levels = c(0,1,2,"3+"),labels = c(0,1,2,"3+"))

train$Self_Employed <- factor(train$Self_Employed, levels = c("No","Yes"),labels = c("No","Yes"))


sapply(train,function(x)sum(is.na(x)))
```

## Filling in the Missing Values
```{r}
train$Gender <- replace(train$Gender,which(is.na(train$Gender)),"Female")

train$Married <- replace(train$Married,which(is.na(train$Married)),"Yes")

train$Dependents <- replace(train$Dependents,which(is.na(train$Dependents)),1)

train$Self_Employed <- replace(train$Self_Employed,which(is.na(train$Self_Employed)),"No")

train$LoanAmount <- replace(train$LoanAmount,which(is.na(train$LoanAmount)),
                            ceiling(mean(c(median(train$LoanAmount,na.rm = TRUE),
                                           mean(train$LoanAmount,na.rm = TRUE)))))

train$Loan_Amount_Term <- replace(train$Loan_Amount_Term,which(is.na(train$Loan_Amount_Term)),360)

train$Credit_History <- replace(train$Credit_History,which(is.na(train$Credit_History)),1)


sapply(train,function(x)sum(is.na(x)))
```

## Creating the Logistic Regression Model
```{r}
training <- train[,-1]
logit <- glm(training$Loan_Status~.,data = training, family = "binomial")

options(scipen=999)      #Prevent Scientific Notation of Numbers
summary(logit)
```

## Generating prediction on training data using the model and checking accuracy of the model
```{r}
pred1 <- predict(logit, training[,-12], type = "response")
confusionMatrix(data=as.factor(ifelse(pred1>0.5,"Y","N")),reference = training$Loan_Status,
                positive = "Y")
```

## Reading the Test Data
```{r}
test <- read.csv("test_lAUu6dG.csv")
```

## Finding the Unique Values
```{r}
sapply(test,function(x)length(unique(x)))
sapply(test[,c(2,3,4,5,6,10,11,12)], function(x)unique(x))
```

## Replacing the Blank Fields with NA
```{r}
sapply(test,function(x)sum(is.na(x)))


test$Gender <- factor(test$Gender,levels = c("Male","Female"), labels = c("Male","Female"))

test$Dependents <- factor(test$Dependents,levels = c(0,1,2,"3+"),labels = c(0,1,2,"3+"))

test$Self_Employed <- factor(test$Self_Employed, levels = c("No","Yes"),labels = c("No","Yes"))


sapply(test,function(x)sum(is.na(x)))
```

## Filling in the Missing Values 
```{r}
test$Gender <- replace(test$Gender,which(is.na(test$Gender)),"Female")

test$Dependents <- replace(test$Dependents,which(is.na(test$Dependents)),1)

test$Self_Employed <- replace(test$Self_Employed,which(is.na(test$Self_Employed)),"No")

test$LoanAmount <- replace(test$LoanAmount,which(is.na(test$LoanAmount)),
                            ceiling(mean(c(median(test$LoanAmount,na.rm = TRUE),
                                           mean(test$LoanAmount,na.rm = TRUE)))))

test$Loan_Amount_Term <- replace(test$Loan_Amount_Term,which(is.na(test$Loan_Amount_Term)),360)

test$Credit_History <- replace(test$Credit_History,which(is.na(test$Credit_History)),1)
```

## Generating Predictions on the Test Data Set
```{r}
pred <- predict(logit, test, type = "response")
test$Loan_Status <- ifelse(pred>0.5,"Y","N")
```

## Showing the Result
```{r}
result <- test[,c(1,13)]
result
```

## Number of Loans Accepted and Rejected
```{r}
data.frame(Accepted=length(which(result$Loan_Status=="Y")),
           Rejected=length(which(result$Loan_Status=="N")))





```

## Exporting the result to csv
```{r}
write.csv(result,"Loan_Prediction_Result.csv",row.names = FALSE)
```

